Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression\thanksref{T1}

### 2018-11-02 revision_1/paper.tex
minor edits
fixed bug in summarizing simulations. some edits to paper

The first example is an analysis of Simon Necomb's 66 measurements of the speed of light; two of which are significant outliers in the lower tail. The full model is a standard location-scale Bayesian model:
\begin{equation}
\beta\sim N(23.6, 2.04^{2}),\  \sigma^{2}\sim IG(5, 10), \ y_{i}\iid N (\beta, \sigma^{2}), i=1,2,\dots, n=66,
\end{equation}
where $y_{i}$ denotes the $i^{th}$ measurement of the passage time of light. $\beta$ is interpreted as the passage time of light with $\sigma^{2}$ representing measurement error.
Four versions of the restricted likelihood are fit with conditioning statistics: 1) Huber's M-estimator for location with Huber's `proposal 2'  for scale 2)  Tukey's M-estimator for location with Huber's `proposal 2'  for scale 3) LMS (least median squares) for location with associated estimator of scale and 4) LTS (least trimmed squares)  for location with associated estimator of scale. Associated tuning parameters for the M-estimators are chosen to achieve $95\%$ efficiency under normality \citep{huber2009} and for comparability, roughly $5\%$ of the residuals are trimmed for LTS.  Additionally, two other common approaches to outlier handling are fit: 1) replacing the normal distribution with a t-distribution and, 2) replacing the normal distribution with a mixture of two normals. The t-model assumes $y_{i}\iid t_{\nu} (\beta, \sigma^{2})$ with $\nu=5$. The prior on $\sigma^{2}$ is $IG(5, \frac{\nu-2}{\nu}10)$ so the prior on the variance is the same as the other models. The mixture takes the form: $y_{i}\iid pN (\beta, \sigma^{2}) + (1-p)N(\beta, 10\sigma^{2})$ assuming the prior $p \sim \text{beta}(20,1)$ on the probability of belonging to the `good' component.

The posteriors of $\beta$ under each model appear in Figure \ref{fig:newcomb_post}. As expected, the posterior under the normal model is pulled downward by the two outliers while the heavy tailed model provides robustness against them. The restricted likelihood methods using the M-estimators and LTS statistics also achieve robustness against the outliers. Conditioning on LMS however, results in a posterior similar to the one under the normal model. The M-estimators provide the most precise posteriors in this case. This is reflected in more precise predictions than the heavy-tailed and mixture model as illustrated by the predictive distributions also shown  in  Figure \ref{fig:newcomb_post}.
\begin{figure}[t]
\centering
{\includegraphics[width = 4in]{figs/speed_of_light_beta.png}}
{\includegraphics[width = 4in]{figs/speed_of_light_predictive.png}}
\caption{Results from the analysis of the speed of light data. Top: Posterior distributions of $\beta$ under each model. Bottom: Log posterior predictive distributions under each model.}
\label{fig:newcomb_post}
\end{figure}

@@

As a second example, a data set measuring the number of telephone calls in Belgium from 1950-1973 is analyzed. The outliers in this case are due to a change in units on which calls were recorded for part of the data set. The full model is a standard normal Bayesian linear regression:
\begin{equation}
{\boldsymbol{\beta}}\sim N_{2}(\boldsymbol{\mu}_{0}, \boldsymbol{\Sigma}_{0}),\  \sigma^{2} \sim IG(a, b), \by \sim N(X\boldsymbol{\beta}, \sigma^{2} I),
\end{equation}
where $\bbeta = (\beta_{0}, \beta_{1})^{\top}$, $\by$ is the vector of the logarithm of the number of calls, and X is the $n\times 2$ design matrix with a vector of 1's in the first column in the year covariate in the second.  Prior parameters are fixed via a fit to the first 3 data points. In particular, $\Sigma_{0} = g\sigma_{0}^2 (X^{\top}X)^{-1}$, with $\sigma_{0} = 0.03$ and $\mu_{0} = (1.87,  0.03)^{\top}$; the MLEs fit to the first three data points. There are $n=21$ remaining data points and the parameter $g$ is set to $21$ reflecting a unit information prior \citep{kass1995reference}. Finally $a = 2$ and $b =1$.

Four models are compared: 1) the normal theory base model 2) A two component normal mixture model, 3) a t-model, and 4) a restricted likelihood model conditioning on Tukey's M-estimator for the slope and intercept with Huber's `proposal 2'  for scale. The mixture model assumes different mean regression functions and variances for each component, but keeps the same, relatively non-informative priors. The probability of belonging to the first component is given a $\text{beta}(5,1)$ prior. The heavy-tailed model fixes the degrees of freedom to 5 with the same adjustment to the prior on $\sigma^{2}$ as above.

The data and  $95\%$ credible bands of the posterior predictive distribution under each model are displayed in Figure \ref{fig:calls_predictive}. The normal theory model is only fit to the obvious non-outlying points. Since the t-model assumes the data are heavy-tailed, the posterior predictive distribution is much wider. On the other hand, the predictive distribution under the restricted likelihood approach is much more precise and is close to that of the normal theory fit that discards the outliers. It is also close to the two component mixture results where the predictive distribution is formulated using only the good component. The mixture model involves explicitly modeling the outlier generating mechanism. In more complex situations where the outlier generating mechanism is transient (i.e. ever changing and more complex than just a unit error in the recording), modeling the outliers becomes more difficult. Like classical robust estimation, the restricted likelihood approach avoids explicitly modeling the outliers. 
\begin{figure}[t]
\centering
{\includegraphics[width = 4in]{figs/calls_predictive.png}}
\caption{Predictive distribution of log(calls) under the Normal theory model fit to the non-outliers, the restricted likelihood model with Tukey's M-estimator for the slope and intercept with Huber's `proposal 2'  for scale, and a heavy-tailed t-distribution model. The first three data points were used to specify the prior with the remaining points used in the posterior fits. See details in the Appendix.}
\label{fig:calls_predictive}
\end{figure}

@@

\section{Simulated Data}
\label{simData}
We study the performance of the restricted likelihood in a hierarchical setting contaminated with outliers. Specifically, simulated data come from the following data generating model:
\begin{align}
\label{gensim2}
\begin{split}
& \theta_{i}  \sim   N(\mu, \tau^{2}),  \ i = 1, 2, \dots, 90  \\ 
& y_{ij} \sim (1-p_{i})N(\theta_{i}, \sigma^{2}) + p_{i}N(\theta_{i}, m_{i}\sigma^{2}),\  j = 1, 2,..., n_{i}
\end{split}
\end{align}
with $\mu = 0, \tau^{2} = 1, \sigma^{2} = 4$. The values of $p_{i}, m_{i}$, and $n_{i}$ depend on the group and are formed using 5 replicates of the full factorial design over factors $p_{i},m_{i},n_{i}$ with levels $p_{i} = .1, .2, .3$, $m_{i} = 9, 25$, and $n_{i} = 25, 50, 100$. This results in 90 groups that have varying levels of outlier contamination and sample size. We wish to build models that offer good prediction for the good portion of data within each group. The full model for fitting is a corresponding normal model without contamination:
\begin{equation}
\label{fullsim2}
\begin{split}
& \mu \propto 1, \  \tau^{2} \propto \tau^{-2}, \\
& \theta_{i}\sim N(\mu, \tau^{2}), \  \sigma^{2}_{i} \sim IG(a_{s}, b_{s}),  \ i = 1, 2, \dots, 90, \\ 
& y_{ij}\sim  N(\theta_{i},\sigma^{2}_{i}), \ j = 1, 2, \dots, n_{i}.
\end{split}
\end{equation}
For the restricted likelihood versions we condition on robust M-estimators of location and scale in each group: $T_{i}(y_{i1}, \dots, y_{in_{i}}) = (\hat\theta_{i}, \hat\sigma^{2}_{i}), i = 1, 2, ..., 90$.  These estimators are solutions to equation \eqref{Mest} (where $x_{i}\equiv 1$) with user specified $\psi$ and $\chi$ functions designed to discount outliers. The two versions use Huber's and Tukey's $\psi$ function, while both versions use Huber's $\chi$ function. The tuning parameters associate with these functions are chosen so that the estimators are $95\%$ efficient under normally distributed data. Both of these are well known M-estimators used in robust regression settings  \citep{huber2009}.  %We will see that his choice has some interesting effects on the results. %These estimators are implemented in the R function \texttt{MASS::rlm} with \cite{huber2009} providing further details.

@@

$K = 30$ data sets are generated from \eqref{gensim2}. For each data set and each pair $(a_{s}, c)$, the Bayesian models are fit using MCMC. The MCMC for the restricted likelihood version requires no further computational details other than those described for the traditional Bayesian model in Section \ref{BayesLinMod}. This is because there are conditioning statistics for each group and the model's conditional independence between the groups allows the data augmentation described earlier to be performed independently within each group. That is, there is a separate Gibbs step for each group  generating group level data matching the statistics for that group.

To assess the predictive capability, the models are compared using Kullback-Leibler (KL) divergence from the distribution of good data to the posterior predictive distribution. Specifically, for the $i^{th}$ group of the $k^{th}$ simulated data set $\by_{k}$ compute:
\begin{equation}
\label{kl}
KL^{(M)}_{ik} = \int \log \frac{f(\tilde y | \theta_{i}, \sigma^{2})}{f_{i}(\tilde y | M, \by_{k})}  f(\tilde y | \theta_{i}, \sigma^{2}) \ dy
\end{equation}
where $M$ indexes the fitting model, $f(\tilde y | \theta_{i}, \sigma^{2}) = N(\tilde y | \theta_{i}, \sigma^{2})$; the mean $\theta_{i}$ and variance $\sigma^{2}$ normal pdf evaluated at $\tilde y$. For the Bayesian models ${f_{i}(\tilde y | M, \by_{k})} = \int f(\tilde y |\theta_{i}, \sigma_{i}^{2}) \pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})d\theta_{i}d\sigma^{2}_{i}$ where $\pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})$ is the posterior for the $i^{th}$ group model parameters under model $M$ for the $k^{th}$ data set. $M$ denotes either the full normal theory model \eqref{fullsim2} or one of the two restricted likelihood versions, along with $a_{s}$ and $c$. For the classical robust fits, we set $f_{i}(\tilde y | M, \by_{k}) = N(\tilde y|\hat\theta_{i}, \hat\sigma^{2}_{i})$ as a groupwise plug-in estimator for the predictive distribution. The classical fits are done separately for each group with no consideration of the hierarchical structure between the groups. The overall mean $\overline{KL}^{(M)}_{{\cdot}{\cdot}} = \frac{1}{90K} \sum_{k = 1}^{K} \sum_{i=1}^{90} KL^{(M)}_{ik}$ is used to compare the models where smaller means correspond to better fits. Sampling variation is summarized with the standard error $SE(\overline{KL}^{(M)}_{{\cdot} k}) = \sqrt{\frac{1}{K(K-1)}\sum_{k = 1}^{K} (\overline{KL}^{(M)}_{{\cdot} k} - \overline{KL}^{(M)}_{{\cdot}{\cdot}})^{2}}$ where $\overline{KL}^{(M)}_{{\cdot} k} = \frac{1}{90}\sum_{i = 1}^{90} KL^{(M)}_{ik}$.

Figure \ref{kl_sim} displays $\overline{KL}^{(M)}_{{\cdot}{\cdot}}$  with error-bars plus/minus one $SE(\overline{KL}^{(M)}_{{\cdot} k})$ for each $a_{s} = 1.25,  5, 10$ and $c = 0.5, 1, 2$. %The panels are for $c = 0.5, 1, 2$ with $a_{s} = 1.25,  5, 10$ on the abscissa. 
The values of $a_{s}$ and $c$, do not effect the classical robust linear models. The normal theory model results are left out as they perform significantly worse. Overall, the results are strikingly in favor of the restricted likelihood methods for the range of hyper-parameter values studied. Undoubtably, the most precise and accurate prior studied is  $c = 1$ and $a_{s} = 10$  and this results in the lowest (and best) average KL for both the Tukey and Huber restricted likelihood versions (shown in the middle panel). $c = 0.5$ performs the worst, but still better than the classical fits. There is evidence that performance starts to degrade in this case when $a_{s} = 10$ as reflected in a larger average KL. Here, the prior mean and precision are $2.22$ and $1.62$ and we suspect this is starting to put too much mass on $\sigma^{2}_{i}$ values much smaller than $\sigma^{2}=4$. For $c = 2$, the performance still improves from $a_{s} = 5$ to $a_{s} = 10$. Here the mean is $8.89$ and precision is only $0.1$; apparently not yet large enough to degrade the performance due to an incorrect mean. Lastly, Tukey's statistic performs marginally better than Huber's. This is likely due to the fact that Tukey's estimator trims extreme outliers completely in the estimation procedure \citep{huber2009}.

\begin{figure}[t]
\centering
\includegraphics[width = 4in]{kl_sim2_facet_scale.png}
\label{kl_sim}
\caption{Average KL - divergence plus/minus one standard error for each value of $a_{s}$ and $c$. The panels correspond to $c = 0.5$ (left), $c=1$ (middle), and $c=2$ (right) with the values of $a_{s}$ on the horizontal axis.}
\end{figure}

It is also interesting to consider the effects of factors $n_{i}$, $p_{i}$, and $m_{i}$. For a given factor and simulation, the $KL^{(M)}_{ik}$  are averaged by factor level. For the Bayesian models, the averages are also taken over the different values of $a_{s}$ and $c$. Figure \ref{kl_mnp} displays these averages for $m,n,$ and $p$ with error bars plus/minus one standard error. The restricted likelihood versions consistently perform better than their classical counterparts. Intuitively, as the amount of contamination ($p$) increases performance degrades as it becomes more difficult to identify the good data. Likewise, as $n$ increases, the performance for the Bayesian methods become closer to that of the their classical counterparts reflecting the diminishing effect of the prior. However, the decrease of KL-diveregence with $m$ and increase with $n$ is somewhat surprising. To investigate, Figure \ref{boxTheta} and \ref{boxSigma} display boxplots of $(\theta_{i} - \hat\theta_{i})$ and $\hat\sigma_{i}$ for each simulated data set. The $\hat\theta_{i}$'s have no systematic bias however the $\hat\sigma_{i}$'s do; consistently biased upward from the value of $\sigma = 2$ for the good portion of data. As $n$ increases, the bias remains the same and the variation in the estimates gets smaller. Thus, the estimates are getting more certain about an incorrect value; explaining the degradation of the KL-divergence. As $m$ increases, the variance in the estimates gets larger and there is a marginal increase in the bias. Thus, for $m=9$ there is more certainty about an incorrect value and  the increased uncertainty at $m=25$ helps to improve the KL-divergence. %Finally, as $p$ grows the bias of the estimates gets significantly larger; resulting in poorer estimation and poorer KL-divergences.

\begin{figure}[t]
\centering
\includegraphics[width = 4in]{kl_sim2_mnp.png}
\caption{Average KL - divergence plus/minus one standard error grouped by the factors $m$ (left), $n$ (middle), and $p$ (right)}
\label{kl_mnp}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width = 4in]{rlm_mnp_theta.png}
\caption{Boxplots $(\theta_{i} - \hat\theta_{i})$ across all simulations separated by the values for $m$ (left), $n$ (middle), $p$ (right) where $\hat\theta_{i}$ are the classical robust estimators (Huber's and Tukey's).}
\label{boxTheta}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width = 4in]{rlm_mnp_sigma.png}
\caption{Boxplots of the classical robust estimators (Huber's and Tukey's) for $\sigma_{i}$ across all simulations separated by the values for $m$ (left), $n$ (middle), $p$ (right). The horizontal line at $\sigma = 2$ highlights the true standard deviation of the `good' data.}
\label{boxSigma}
\end{figure}

This simulation shows the potential of the restricted likelihood while highlighting some interesting observations. Specifically, the choice of summary statistics, along with corresponding tuning parameters is important. For the tuning parameters, the standard choice of $95\%$ efficiency at the normal was used.  Under the data generating model here, this choice results in bias in the scale estimation which affects the performance of the method. These choices must be made in both the classical and Bayesian settings. The Bayesian setting allows for the incorporation of informative prior information which, as shown in this example, can dramatically improve prediction. The results are sensitive to the prior, but here we have observed good relative improvement over the classical counterparts for a range of prior choices.

@@

The data a grouped by states with a varying number of agencies by state. Identifiers such as agency/agent names are removed. Likewise, state labels and agency types (identifying the varying contractual agreements) have been made generic to protect the proprietary nature of the data.

@@

The presence of numerous outliers in the data implies that both training and validation data will contain 
outliers.  For this reason, the evaluation must be robust to a certain fraction of bad data.  
The two main strategies are to robustify the evaluation function \citep[e.g.,][]{ronchetti1997} or 
to retain the desired evaluation function and trim cases \citep{jung2014}.  Here,
we pursue the trimming approach with log predictive density for the Bayesian models and log plug-in 
maximum likelihood for the classical fits used as the evaluation function.

>>>>

The first example is an analysis of Simon Necomb's 66 measurements of the passage time of light \citep{stigler1977}; two of which are significant outliers in the lower tail. The full model is a standard location-scale Bayesian model also used in \cite{lee2014}:
\begin{equation}
\beta\sim N(23.6, 2.04^{2}),\  \sigma^{2}\sim IG(5, 10), \ y_{i}\iid N (\beta, \sigma^{2}), i=1,2,\dots, n=66,
\end{equation}
where $y_{i}$ denotes the $i^{th}$ (recoded) measurement of the passage time of light. $\beta$ is interpreted as the passage time of light with the deviations $y_i - \beta_i$ representing measurement error.
Four versions of the restricted likelihood are fit with conditioning statistics: 1) Huber's M-estimator for location with Huber's `proposal 2'  for scale 2)  Tukey's M-estimator for location with Huber's `proposal 2'  for scale 3) LMS (least median squares) for location with associated estimator of scale and 4) LTS (least trimmed squares)  for location with associated estimator of scale. The tuning parameters for the M-estimators are chosen to achieve $95\%$ efficiency under normality \citep{huber2009} and, for comparability, roughly $5\%$ of the residuals are trimmed for LTS.  Two additional approaches to outlier handling are fit: 1) the normal distribution is replaced with a t-distribution and, 2) the normal distribution is replaced with a mixture of two normals. The t-model assumes $y_{i}\iid t_{\nu} (\beta, \sigma^{2})$ with $\nu=5$. The prior on $\sigma^{2}$ is $IG(5, \frac{\nu-2}{\nu}10)$ ensures that the prior on the variance is the same as the other models. The mixture takes the form: $y_{i}\iid pN (\beta, \sigma^{2}) + (1-p)N(\beta, 10\sigma^{2})$ with the prior $p \sim \text{beta}(20,1)$ on the probability of belonging to the `good' component.

%The posteriors of $\beta$ under each model appear in Figure \ref{fig:newcomb_post}. As expected, the posterior under the normal model is pulled downward by the two outliers while the heavy-tailed model provides robustness against them. The restricted likelihood methods using the M-estimators and LTS statistics also achieve robustness against the outliers. Conditioning on LMS however, results in a posterior similar to the one under the normal model. The M-estimators provide the most precise posteriors in this case. This is reflected in more precise predictions than the heavy-tailed and mixture model as illustrated by the predictive distributions also shown  in  Figure \ref{fig:newcomb_post}.
The posterior of $\beta$ under each model appears in Figure \ref{fig:newcomb_post}.  The posteriors group into two batches.  The normal model and restricted likelihood with LMS do not discount the outliers and have posteriors centered at low values of $\beta$.  These posteriors are also quite diffuse.  In contrast, the t model and the other restricted likelihood methods discount the outliers and have posterior centered at higher values.  There is modest variation among these centers.  Posteriors in this second group have less dispersion than those in the first group.   
\begin{figure}[t]
\centering
{\includegraphics[width = 4in]{figs/speed_of_light_beta.png}}
{\includegraphics[width = 4in]{figs/speed_of_light_predictive.png}}
\caption{Results from the analysis of the speed of light data. Top: Posterior distributions of $\beta$ under each model. Bottom: Log posterior predictive distributions under each model. The differences in the tails are emphasized in the bottom plot. The horizontal axis is strategically labeled to help compare the centers of the distributions in each of the plots.}
\label{fig:newcomb_post}
\end{figure}

The pattern for predictive distributions differs (see bottom plot in Figure \ref{fig:newcomb__post}).  The normal and t models have widely dispersed predictive distributions.  The other predictive distributions show much greater concentration.  The restricted likelihood fits based on M-estimators (Tukey's and Huber's) are centered appropriately and are concentrated.  The LMS predictive is concentrated, but it is poorly centered.

Overall, we find that the restricted likelihood methods based on M-estimators provide the most attractive analysis for these data.  They provide sharp and appropriate inference for parameters ($\beta$) and for prediction.

@@

As a second example, a data set measuring the number of telephone calls in Belgium from 1950-1973 is analyzed. The outliers in this case are due to a change in measurement units on which calls were recorded for part of the data set. Specifically, for years 1964-1969 and parts of 1963 and 1970, the length of calls in minutes were recorded rather than the total number \citep{rousseeuw1987}. The full model is a standard normal Bayesian linear regression:
\begin{equation}
{\boldsymbol{\beta}}\sim N_{2}(\boldsymbol{\mu}_{0}, \boldsymbol{\Sigma}_{0}),\  \sigma^{2} \sim IG(a, b), \by \sim N(X\boldsymbol{\beta}, \sigma^{2} I),
\end{equation}
where $\bbeta = (\beta_{0}, \beta_{1})^{\top}$, $\by$ is the vector of the logarithm of the number of calls, and X is the $n\times 2$ design matrix with a vector of 1's in the first column in the year covariate in the second.  Prior parameters are fixed via a fit to the first 3 data points. In particular, the prior covariance for $\beta$ is set to $\Sigma_{0} = g\sigma_{0}^2 (X_{p}^{\top}X_{p})^{-1}$, with $X_{p}$ the $3\times 2$ design matrix for the first $3$ data points, $g=n=21$, and $\sigma_{0} = 0.03$ and $\mu_{0} = (1.87,  0.03)^{\top}$; the MLEs fit to the first three data points. This has the spirit of a unit information prior \citep{kass1995reference} but uses a design matrix for prior data not used in the fit. Finally $a = 2$ and $b =1$.

Four models are compared: 1) the normal theory base model 2) A two component normal mixture model, 3) a t-model, and 4) a restricted likelihood model conditioning on Tukey's M-estimator for the slope and intercept with Huber's `proposal 2'  for scale. Each model is fit to the remaining 21 data points. The normal theory model is also fit a second time after removing observations 14-21 (years 1963 - 1970). This includes the obvious large outliers as well as the two smaller outliers at the beginning and end of this sequence of points caused by the change in measurement units. The mixture model assumes different mean regression functions and variances for each component, but keeps the same, relatively non-informative priors. The probability of belonging to the first component is given a $\text{beta}(5,1)$ prior. The heavy-tailed model fixes the degrees of freedom to 5 with the same adjustment to the prior on $\sigma^{2}$ as above.

The data and  $95\%$ credible bands of the posterior predictive distribution under each model are displayed in Figure \ref{fig:calls_predictive}. The normal model fit to all cases results in a very wide posterior predictive distribution due to an inflated estimate of the variance. Since the t-model assumes the data are heavy-tailed, the posterior predictive distribution is also wide. Additionally, both of these are biased upward by the outliers in the upper tail of the assumed data distribution. On the other hand, the predictive distribution under the restricted likelihood approach is much more precise and is close to that of the normal theory fit to the non-outlying cases. It is also close to the two component mixture results where the predictive distribution is formulated using only the good component. The mixture model involves explicitly modeling the outlier generating mechanism. In more complex situations where the outlier generating mechanism is transient (i.e. ever changing and more complex than just a unit error in the recording), modeling the outliers becomes more difficult. As in classical robust estimation, the restricted likelihood approach avoids explicitly modeling the outliers. 
\begin{figure}[t]
\centering
{\includegraphics[width = 4in]{figs/calls_predictive.png}}
\caption{Pointwise posterior predictive intervals of log(calls) under the normal theory model fit to the non-outliers, the restricted likelihood model with Tukey's M-estimator for the slope and intercept with Huber's `proposal 2'  for scale, and a heavy-tailed t-distribution model. The first three data points were used to specify the prior with each model using the remaining 21 for fitting. The normal theory model was also fit after removing observations 14-20 (years 1963 - 1970).}
\label{fig:calls_predictive}
\end{figure}
%{\bf For the figure, add fit of normal model to all cases}

@@

\section{Simulated Data}
\label{simData}
We study the performance of restricted likelihood methods in a hierarchical setting where the data are contaminated with outliers. Specifically, simulated data come from the following model:\begin{align}
\label{gensim2}
\begin{split}
& \theta_{i}  \sim   N(\mu, \tau^{2}),  \ i = 1, 2, \dots, 90  \\ 
& y_{ij} \sim (1-p_{i})N(\theta_{i}, \sigma^{2}) + p_{i}N(\theta_{i}, m_{i}\sigma^{2}),\  j = 1, 2,..., n_{i}
\end{split}
\end{align}
with $\mu = 0, \tau^{2} = 1, \sigma^{2} = 4$. The values of $p_{i}, m_{i}$, and $n_{i}$ depend on the group and are formed using 5 replicates of the full factorial design over factors $p_{i},m_{i},n_{i}$ with levels $p_{i} = .1, .2, .3$, $m_{i} = 9, 25$, and $n_{i} = 25, 50, 100$. This results in 90 groups that have varying levels of outlier contamination and sample size. We wish to build models that offer good prediction for the good portion of data within each group. The full model for fitting is a corresponding normal model without contamination:
\begin{equation}
\label{fullsim2}
\begin{split}
& \mu \propto 1, \  \tau^{2} \propto \tau^{-2}, \\
& \theta_{i}\sim N(\mu, \tau^{2}), \  \sigma^{2}_{i} \sim IG(a_{s}, b_{s}),  \ i = 1, 2, \dots, 90, \\ 
& y_{ij}\sim  N(\theta_{i},\sigma^{2}_{i}), \ j = 1, 2, \dots, n_{i}.
\end{split}
\end{equation}
For the restricted likelihood versions we condition on robust M-estimators of location and scale in each group: $T_{i}(y_{i1}, \dots, y_{in_{i}}) = (\hat\theta_{i}, \hat\sigma^{2}_{i}), i = 1, 2, ..., 90$.  These estimators are solutions to equation \eqref{Mest} (where $x_{i}\equiv 1$) with user specified $\psi$ and $\chi$ functions designed to discount outliers. The two versions use Huber's and Tukey's $\psi$ function, while both versions use Huber's $\chi$ function. The tuning parameters associated with these functions are chosen so that the estimators are $95\%$ efficient under normally distributed data. These classical M-estimators are commonly used in robust regression settings  \citep{huber2009}.  %We will see that his choice has some interesting effects on the results. %These estimators are implemented in the R function \texttt{MASS::rlm} with \cite{huber2009} providing further details.

@@

$K = 30$ data sets are generated from \eqref{gensim2}. For each data set and each pair $(a_{s}, c)$, the Bayesian models are fit using MCMC. The MCMC for the restricted likelihood version requires no computational details other than those described for the traditional Bayesian model in Section \ref{BayesLinMod}. This is because there are conditioning statistics for each group and the model's conditional independence between the groups allows the data augmentation described earlier to be performed independently within each group. That is, there is a separate Gibbs step for each group to generate the group level data matching the statistics for that group.

To assess predictive capability, the models are compared using Kullback-Leibler (KL) divergence from the distribution of good data to the posterior predictive distribution. Specifically, for the $i^{th}$ group of the $k^{th}$ simulated data set $\by_{k}$ compute:
\begin{equation}
\label{kl}
KL^{(M)}_{ik} = \int \log \frac{f(\tilde y | \theta_{i}, \sigma^{2})}{f_{i}(\tilde y | M, \by_{k})}  f(\tilde y | \theta_{i}, \sigma^{2}) \ dy
\end{equation}
where $M$ indexes the fitting model, $f(\tilde y | \theta_{i}, \sigma^{2}) = N(\tilde y | \theta_{i}, \sigma^{2})$; the normal density function with mean $\theta_{i}$ and variance $\sigma^{2}$, evaluated at $\tilde y$. For the Bayesian models ${f_{i}(\tilde y | M, \by_{k})} = \int f(\tilde y |\theta_{i}, \sigma_{i}^{2}) \pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})d\theta_{i}d\sigma^{2}_{i}$ where $\pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})$ is the posterior for the $i^{th}$ group model parameters under model $M$ for the $k^{th}$ data set. $M$ denotes either the full normal theory model \eqref{fullsim2} or one of the two restricted likelihood versions, along with $a_{s}$ and $c$. For the classical robust fits, we set $f_{i}(\tilde y | M, \by_{k}) = N(\tilde y|\hat\theta_{i}, \hat\sigma^{2}_{i})$ as a groupwise plug-in estimator for the predictive distribution. The classical fits are done separately for each group with no consideration of the hierarchical structure between the groups. The overall mean $\overline{KL}^{(M)}_{{\cdot}{\cdot}} = \frac{1}{90K} \sum_{k = 1}^{K} \sum_{i=1}^{90} KL^{(M)}_{ik}$ is used to compare the models, where smaller means correspond to better fits. Sampling variation is summarized with the standard error between the $K=30$ replicates in the simulation:  $SE(\overline{KL}^{(M)}_{{\cdot} k}) = \sqrt{\frac{1}{K(K-1)}\sum_{k = 1}^{K} (\overline{KL}^{(M)}_{{\cdot} k} - \overline{KL}^{(M)}_{{\cdot}{\cdot}})^{2}}$ where $\overline{KL}^{(M)}_{{\cdot} k} = \frac{1}{90}\sum_{i = 1}^{90} KL^{(M)}_{ik}$.

Figure \ref{kl_sim} displays $\overline{KL}^{(M)}_{{\cdot}{\cdot}}$  with error-bars plus/minus one $SE(\overline{KL}^{(M)}_{{\cdot} k})$ for each $a_{s} = 1.25,  5, 10$ and $c = 0.5, 1, 2$. %The panels are for $c = 0.5, 1, 2$ with $a_{s} = 1.25,  5, 10$ on the abscissa. 
The values of $a_{s}$ and $c$, do not effect the classical robust linear models. The average KL  for the normal theory model ranges between $0.31$ and $0.37$ which is much worse than the robust methods and hence these are left out of the figure. Overall, the results are strikingly in favor of the restricted likelihood methods for the range of hyper-parameter values studied. Undoubtably, the most precise and accurate prior studied is  $c = 1$ and $a_{s} = 10$  and this results in the lowest (and best) average KL for both the Tukey and Huber restricted likelihood versions (shown in the middle panel). $c = 0.5$ performs the worst, but still better than the classical fits. There is evidence that performance starts to degrade in this case when $a_{s} = 10$ as reflected in a larger average KL. Here, the prior mean and precision are $2.22$ and $1.62$ and we suspect this is starting to put too much mass on $\sigma^{2}_{i}$ values much smaller than $\sigma^{2}=4$. For $c = 2$, the performance still improves from $a_{s} = 5$ to $a_{s} = 10$. Here the mean is $8.89$ and precision is only $0.1$; apparently not yet large enough to degrade the performance due to an incorrect mean. Lastly, Tukey's statistic performs marginally better than Huber's. This is likely due to the fact that Tukey's estimator trims extreme outliers completely in the estimation procedure \citep{huber2009}.

\begin{figure}[t]
\centering
\includegraphics[width = 4in]{kl_sim2_facet_scale.png}
\label{kl_sim}
\caption{Average KL - divergence plus/minus one standard error for each value of $a_{s}$ and $c$ ($\overline{KL}^{(M)}_{{\cdot}{\cdot}} \pm SE(\overline{KL}^{(M)}_{{\cdot} k})$). Smaller values represent better fits. The panels correspond to $c = 0.5$ (left), $c=1$ (middle), and $c=2$ (right) with the values of $a_{s}$ on the horizontal axis. The average KL  for the normal theory model ranges between $0.31$ and $0.37$ and are left out of the figure.}
\end{figure}

%% DELETE?
%It is also interesting to consider the effects of factors $n_{i}$, $p_{i}$, and $m_{i}$. For a given factor and simulation, the $KL^{(M)}_{ik}$  are averaged by factor level. For the Bayesian models, the averages are also taken over the different values of $a_{s}$ and $c$. Figure \ref{kl_mnp} displays these averages for $m,n,$ and $p$ with error bars plus/minus one standard error. The restricted likelihood versions consistently perform better than their classical counterparts. Intuitively, as the amount of contamination ($p$) increases performance degrades as it becomes more difficult to identify the good data. Likewise, as $n$ increases, the performance for the Bayesian methods become closer to that of the their classical counterparts reflecting the diminishing effect of the prior. However, the decrease of KL-diveregence with $m$ and increase with $n$ is somewhat surprising. To investigate, Figure \ref{boxTheta} and \ref{boxSigma} display boxplots of $(\theta_{i} - \hat\theta_{i})$ and $\hat\sigma_{i}$ for each simulated data set. The $\hat\theta_{i}$'s have no systematic bias however the $\hat\sigma_{i}$'s do; consistently biased upward from the value of $\sigma = 2$ for the good portion of data. As $n$ increases, the bias remains the same and the variation in the estimates gets smaller. Thus, the estimates are getting more certain about an incorrect value; explaining the degradation of the KL-divergence. As $m$ increases, the variance in the estimates gets larger and there is a marginal increase in the bias. Thus, for $m=9$ there is more certainty about an incorrect value and  the increased uncertainty at $m=25$ helps to improve the KL-divergence. %Finally, as $p$ grows the bias of the estimates gets significantly larger; resulting in poorer estimation and poorer KL-divergences.
% 
%\begin{figure}[t]
%\centering
%\includegraphics[width = 4in]{kl_sim2_mnp.png}
%\caption{Average KL - divergence plus/minus one standard error grouped by the factors $m$ (left), $n$ (middle), and $p$ (right)}
%\label{kl_mnp}
%\end{figure}
%\begin{figure}[t]
%\centering
%\includegraphics[width = 4in]{rlm_mnp_theta.png}
%\caption{Boxplots $(\theta_{i} - \hat\theta_{i})$ across all simulations separated by the values for $m$ (left), $n$ (middle), $p$ (right) where $\hat\theta_{i}$ are the classical robust estimators (Huber's and Tukey's).}
%\label{boxTheta}
%\end{figure}
%\begin{figure}[t]
%\centering
%\includegraphics[width = 4in]{rlm_mnp_sigma.png}
%\caption{Boxplots of the classical robust estimators (Huber's and Tukey's) for $\sigma_{i}$ across all simulations separated by the values for $m$ (left), $n$ (middle), $p$ (right). The horizontal line at $\sigma = 2$ highlights the true standard deviation of the `good' data.}
%\label{boxSigma}
%\end{figure}
%% --------
{EDIT THIS} This simulation shows the potential of the restricted likelihood while highlighting some interesting observations. Specifically, the choice of summary statistics, along with corresponding tuning parameters is important. For the tuning parameters, the standard choice of $95\%$ efficiency at the normal was used.  Under the data generating model here, this choice results in bias in the scale estimation which affects the performance of the method. These choices must be made in both the classical and Bayesian settings. The Bayesian setting allows for the incorporation of informative prior information which, as shown in this example, can dramatically improve prediction. The results are sensitive to the prior, but here we have observed good relative improvement over the classical counterparts for a range of prior choices.

@@

The data are grouped by states with a varying number of agencies by state. Identifiers such as agency/agent names are removed. Likewise, state labels and agency types (identifying the varying contractual agreements) have been made generic to protect the proprietary nature of the data.

@@

The presence of numerous outliers in the data implies that both training and validation data will contain 
outliers.  For this reason, the evaluation must be robust to a certain fraction of bad data.  
The two main strategies are to robustify the evaluation function \citep[e.g.,][]{ronchetti1997} or 
to retain the desired evaluation function and trim cases \citep{jung2014}.  Here,
we pursue the trimming approach with log predictive density for the Bayesian models and log density from plug-in 
maximum likelihood for the classical fits used as the evaluation function.
